Incidence of confirmed Covid-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)

### 2020-07-05 manuscript/manuscript.Rmd
manuscript version 0.4

---
title: "Incidence of confirmed COVID-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  word_document:
    toc: yes
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

residents_episodes <- tline %>% 
  filter(rday==1) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

@@

@@

km_ct_residents_deaths <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    deaths = sum(new_deaths),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = deaths, 
                  pop = occupancy, subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "COVID-19-related death"
  )

@@

desc_residents$DAT_CFR_RR <- relative_risk(desc_residents$DAT_CFR$y, desc_residents$DAT_CFR$n - desc_residents$DAT_CFR$y, desc_residents$DAT_BFR$y, desc_residents$DAT_BFR$n - desc_residents$DAT_BFR$y)

desc_residents$DAT_CFR_RR_SE <- relative_risk_SE(desc_residents$DAT_CFR$y, desc_residents$DAT_CFR$n - desc_residents$DAT_CFR$y, desc_residents$DAT_BFR$y, desc_residents$DAT_BFR$n - desc_residents$DAT_BFR$y)

desc_residents[paste0(
  "DAT_CFR", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_residents$DAT_CFR$y,
  desc_residents$DAT_CFR$n
)[c("proportion", "lower", "upper")] *100

desc_residents[paste0(
  "DAT_BFR", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_residents$DAT_BFR$y,
  desc_residents$DAT_BFR$n
)[c("proportion", "lower", "upper")] *100

@@

## OUTBREAK HOMES 
desc_residents_outbreak_only <- list()
desc_residents_outbreak_only$outbreak_homes <- unique(
  home_total_confirmed$home_code[home_total_confirmed$CT_confirmed != 0],
  distinct(filter(new_cases, tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0), home_code)$home_code
)

@@

```{r captions}
## TABLES
table_nums <- captioner::captioner(prefix = "Table")
tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, contract beds, and residents within FSHCG and nationwide")
tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections in residents according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections among staff according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))

@@

@@

**Background** COVID-19 has caused substantial morbidity and mortality in care home residents and staff. National surveys suggest infections are declining in the general population in the UK, but outbreaks persist in care homes. We measured the incidence of infections among care home staff and residents, using administrative data from the Four Seasons Health Care Group (FSHCG), a national care home chain representing 9% of all beds in England, Scotland and Northern Ireland.

**Methods:** Retrospective cohort study using individual-level electronic records from residents (incidents reports) and daily care home counts of infection for both residents and staff. We estimated the incidence proportion and incidence rate of symptoms, confirmed infections, and deaths in residents and staff. The rate of confirmed infections was compared with data from a national community survey. We tested a range of risk factors for resident infection and mortality in multivariate Cox proportional hazards models.

**Findings:** An estimated `r formatbm(desc_residents$CT_total_residents)` residents were present in `r dplyr::n_distinct(residents$home_code)` FSHCG homes between `r format(date_min_tallies, "%e %B %Y")` and `r format(date_max_tallies, "%e %B %Y")`. During the period `r formatbm(desc_residents$CT_total_symptomatic)` residents developed COVID-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), and `r formatbm(desc_residents$CT_total_confirmed)` (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) had an infection confirmed with nasopharyngeal RNA tests. The confirmed infection incidence rate was `r tbrounding(desc_residents$CT_IR_confirmed)` per 100,000 resident-days [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] across all care homes, including `r desc_residents$CT_zero_case` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) which had not reported a single confirmed infection or death.
Within England, the confirmed infection rate was `r tbrounding(england_results$results$measure[2,"estimate"])` times [`r tbrounding(england_results$results$measure[2,"lower"]) `; `r tbrounding(england_results$results$measure[2,"upper"])`] greater in care homes than in community households. Out of `r formatbm(desc_staff$total_staff)` staff, `r formatbm(desc_staff$CT_confirmed_total)` confirmed infections were counted (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]), corresponding to a rate of incidence of 
`r tbrounding(desc_staff$CT_confirmed_IR)` per 100,000 staff-days [`r tbrounding(desc_staff$CT_symptomatic_IR_lower)`; `r tbrounding(desc_staff$CT_symptomatic_IR_upper)`].

Several risk factors were independently associated with residents' confirmed infection hazards.
Males and those aged 85 years had significantly greater hazards of confirmed infection. Homes with a ratio of occupants to bedrooms (measured at the beginning of the epidemic) in excess of 1.00 were found to have confirmed infections hazards 10 times [7·0; 15] greater than homes with a ratio of 0.70 to 0.85. Conversely, an increase by 10 percentage points in the staff-to-bed ratio was associated with a 23% reduction [18%, 29%] in the hazards of confirmed infections.

All-cause deaths both within homes and in hospitals were recorded for occupants of private beds. 
The all-cause mortality proportion was 
`r tbrounding(desc_residents$DAT_p_death)`% [`r tbrounding(desc_residents$DAT_p_death_lower)`%; `r tbrounding(desc_residents$DAT_p_death_upper)`%], with a confirmed infection case-fatality rate of
`r tbrounding(desc_residents$DAT_CFR)`% [`r tbrounding(desc_residents$DAT_CFR_lower)`%; `r tbrounding(desc_residents$DAT_CFR_upper)`%]. All-cause mortality was strongly associated with both older age and being male, but not with rates of staffing or occupancy. Adjusted mortality hazard ratios for asymptomatic and symptomatic confirmed cases respectively were 2·1 [1·3; 3·6] and 8·5 [7·1; 10] compared with asymptomatic residents who never tested positive for SARS-CoV-2.

**Interpretation** Despite uncertain case ascertainment in care homes, internal surveillance suggests a greater transmission than in the community. Organisational factors such as staffing and occupancy are strongly associated with rates of confirmed infections and can consequently place residents at increased risks of death. The absence of high-quality case ascertainment across the social care sector hinders local infection prevention initiatives.

@@

Care home residents are vulnerable to infections due to their advanced age and high prevalence of comorbidity. This is compounded by frequent exposure to infections, for example through contact with infected staff members. At the peak of the pandemic the number of deaths in UK care home residents was threefold higher than the equivalent period in 2019.[@onsDR2020]

Information on disease incidence and prevalence, and how they vary across care homes is essential to inform the public health response to future waves of infection, and in particular decisions around the need for widespread universal testing in care home staff and residents. Yet, estimates of infection incidence and prevalence over this time period are lacking due to delays in establishing universal testing for care home residents and staff.

For example, at the start of the pandemic, testing for SARS-CoV-2 infection testing using reverse transcriptase polymerase chain reaction (RT-PCR) was only routinely available for care home residents or staff who were admitted to hospital. Testing was also performed in care homes that notified an outbreak to Public Health England, but this was limited to a maximum of five test per care home. Widespread testing for residents only became available on 11 May 2020 through the whole care home testing programme.[@dhscTesting2020] As a result, national estimates of incidence and prevalence based on test results obtained prior to 11 May substantially underestimate the burden of infection in care home residents.

Outbreak investigations constitute an important source of epidemiological data on the burden of disease in care homes.[@McMichael2020; @Dora2020] A living systematic review[@Salcher-Konrad2020] of outbreak investigations found incidence proportions ranging between 0% and 72% among residents and between 2% and 64% among staff of long term care facilities. Despite this, estimates of this nature cannot be generalised due to the unknown proportion of care homes that have not reported any outbreaks.

This study obtained data from `r formatbm(dplyr::n_distinct(residents$home_code))` care homes from the Four Seasons Health Care Group (FSHCG), one of the UK's largest independent provider of long-term and respite residential and nursing care. It estimated incidence of SARS-CoV-2 symptoms, diagnoses and all-cause mortality across all homes. In addition, it tested and identified individual and organisational risk factors associated with the rate of confirmed infections and deaths.

@@

In 2020, FSHCG counted 9,568 beds, representing 9% of all beds in England, Scotland and Northern Ireland (`r table_nums("tab_national_descriptives", display = "cite")`). Under 9% of these beds were contracted to local authorities for their statutory social care responsibilities.

FSHCG collects administrative data on their private residents through a variety of electronic systems that are primarily used for billing purposes. This includes individual-level data on the dates of care home entry and exit for residents, demographic information, and reports of infection through an incidents management system known as "Datix". Such records do not include residents occupying beds contracted by local authorities as part of block agreements, but home-level aggregate information was nevertheless available on all residents (`r fig_nums("fig_study_diagram", display = "cite")`).

@@

On 24 March 2020, FSHCG introduced a bespoke reporting system to monitor COVID-19 cases in each care home. This new system requires care homes to report daily tallies of new cases identified among staff and residents to their FSHCG regional manager. Daily counts were obtained for every care home:

- new symptomatic cases (residents)
- newly confirmed infection in home (residents)
- newly confirmed infection in hospital (residents)
- deaths related to COVID-19 (residents previously confirmed or with a death attributed to SARS-CoV-2 by coroner)
- symptomatic cases (staff)
- newly confirmed cases (staff)

In addition, the number of beds currently occupied was reported every Sunday. This weekly information was used to approximate daily occupancy using linear interpolation.

These aggregate data related to private residents, residents occupying local authority contract beds, as well as a minority of residents receiving a different category of support (eg. young people) whose individual-level records were not extracted.

@@

- **a residents dataset** containing one record per resident per care home, alongside residents' gender, date of birth, most recent dates of admission and discharge, date of first admission, type of stay (residential/nursing) and care (general, dementia, elderly). This dataset does not include any records for occupants of local authority 'block contract' beds  ($n=$ `r formatbm(sum(desc_homes[["Total contract beds"]], na.rm=T))`).
- **an incidents dataset** containing `r formatbm(desc_before_exclusions$nrow_incidents)` reports filed by care home staff into FSHCG's Datix incidents management system. Mandated report field were: resident forename, surname, care home identifier, incident date/time, and date/time of reporting. In addition, reports could record: date of birth, gender, information on COVID-19 symptoms, tests and test results, resident current location (home/hospital), and death.

>>>>

---
title: "Incidence of confirmed Covid-19 cases in care home staff and residents: Cohort study using electronic records in a large UK care home provider (March-June 2020)"
# author: "Dutey-Magni, PF, Williams, H, Hayward, A, Shallcross, L"
bibliography: references.bib
csl: elsevier-vancouver.csl
output:
  word_document:
    toc: yes
    reference_docx: template.docx
  html_document:
    df_print: paged
    toc: yes
    code_folding: hide
link-citations: yes
always_allow_html: yes
---

@@

@@

home_first_case <- bind_rows(
  new_cases %>% 
    dplyr::filter(tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(date),
                     source = "CT"),
  residents_overview %>% 
    dplyr::filter(confirmed == 1) %>% 
    dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(covid_first_confirmed, na.rm = T), source="DAT")
) %>% 
  dplyr::group_by(home_code) %>% 
    dplyr::summarise(first_case_date = min(first_case_date)) %>% 
  tidyr::crossing(tibble(outbreak = c(0, 1)))

residents_episodes <- tline %>% 
  filter(rday==1)  %>% 
  left_join(home_first_case) %>%
  mutate(outbreak = dplyr::case_when(
    date >= first_case_date ~ 1L,
    date < first_case_date ~ 0L,
    TRUE ~ 0L
  )) %>% 
  group_by(resident_id, home_code, dob, gender, 
           status, home_type, user_type2,
           susceptible_symptomatic, 
           susceptible_confirmed,
           outbreak) %>% 
  summarise(entry_date = min(date),
            exit_date = max(date+1),
            resident_days = sum(rday),
            susceptible_symptomatic_days = sum(susceptible_symptomatic),
            susceptible_confirmed_days = sum(susceptible_confirmed)) %>% 
  mutate(date1 = as.numeric( entry_date - date_min),
         date2 = as.numeric(exit_date - date_min),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95)) %>% 
  mutate(symptomatic = -susceptible_symptomatic+1,
         confirmed = -susceptible_confirmed+1,
         symptomatic_confirmed = (-susceptible_symptomatic+1)*(-susceptible_confirmed+1),
         date2 = if_else(date1 == date2, date2 + 1, date2)) %>% 
  mutate(covid_status = case_when(
    symptomatic==0 & confirmed == 0 ~ "Negative",
    symptomatic==1 & confirmed == 0 ~ "Symptomatic (not confirmed)",
    symptomatic==0 & confirmed == 1 ~ "Asymptomatic confirmed",
    symptomatic==1 & confirmed == 1 ~ "Symptomatic confirmed"
  )) %>% 
  mutate(covid_status = factor(covid_status, 
                               levels = c("Negative", "Symptomatic (not confirmed)",
                                          "Asymptomatic confirmed", "Symptomatic confirmed")),
         ageg = cut_age_denary(compute_age(dob, date_min), age.min = 75, age.max = 95))   %>% 
  left_join(select(reference_homes, home_code, country, region_UK,
                   staff_to_bed_ratio, carer_to_bed_ratio, bed_tot_category, 
                   beds, pct_occupancy, no_bedrooms, 
                   occ_bedroom_ratio, occ_bedroom_ratio_g5, occ_bedroom_ratio_g6)) %>% 
  group_by(resident_id) %>% 
  mutate(death_status = if_else(
    date1 == max(date1),
    as.integer(status == "Deceased"),
    0L
  )) %>% 
  ungroup()

@@

km_ct_residents_deaths <- residents_ct %>% 
  group_by(date) %>% 
  summarise(
    deaths = sum(new_deaths),
    occupancy = sum(occupancy)
  ) %>% 
  km_ct_estimator(df = ., t = date, d = deaths, 
                  pop = occupancy, subtract_cases = FALSE, overwrite = TRUE) %>% 
  ungroup() %>% 
  mutate(
    St_max = S_t^exp(1.96*S_t_SElog),
    St_min = S_t^exp(-1.96*S_t_SElog),
    event = "Covid-19-related death"
  )

@@

desc_residents$DAT_CFR_RR <- relative_risk(
  desc_residents$DAT_CFR$y, 
  desc_residents$DAT_CFR$n - desc_residents$DAT_CFR$y, 
  desc_residents$DAT_BFR$y, 
  desc_residents$DAT_BFR$n - desc_residents$DAT_BFR$y)

desc_residents$DAT_CFR_RR_SE <- relative_risk_SE(
  desc_residents$DAT_CFR$y, 
  desc_residents$DAT_CFR$n - desc_residents$DAT_CFR$y, 
  desc_residents$DAT_BFR$y, 
  desc_residents$DAT_BFR$n - desc_residents$DAT_BFR$y)

desc_residents$DAT_CFR_RR_lower <- exp(log(desc_residents$DAT_CFR_RR) - 1.96 * desc_residents$DAT_CFR_RR_SE)
desc_residents$DAT_CFR_RR_upper <- exp(log(desc_residents$DAT_CFR_RR) + 1.96 * desc_residents$DAT_CFR_RR_SE)

desc_residents[paste0(
  "DAT_CFR_rate", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_residents$DAT_CFR$y,
  desc_residents$DAT_CFR$n
)[c("proportion", "lower", "upper")] *100

desc_residents[paste0(
  "DAT_BFR_rate", c("", "_lower", "_upper")
)] <- epitools::binom.exact(
  desc_residents$DAT_BFR$y,
  desc_residents$DAT_BFR$n
)[c("proportion", "lower", "upper")] *100

@@

## OUTBREAK HOMES 
desc_residents_outbreak_only <- list()
desc_residents_outbreak_only$outbreak_homes <- unique(
  c(home_total_confirmed$home_code[home_total_confirmed$CT_confirmed != 0],
  distinct(filter(new_cases, tidyr::replace_na(new_deaths_hospital, 0) > 0 |
                    tidyr::replace_na(new_deaths_home, 0)>0), home_code)$home_code,
  distinct(filter(residents_overview, confirmed == 1), home_code)$home_code)
)

desc_residents$DAT_BFR_outbreak_homes <- residents_overview %>%
  filter(home_code %in% desc_residents_outbreak_only$outbreak_homes & confirmed != 1) %>%
  summarise(n = n(),
            y = sum(status == "Deceased"))

desc_residents$DAT_BFR_other_homes <- residents_overview %>%
  filter((!home_code %in% desc_residents_outbreak_only$outbreak_homes) & confirmed != 1) %>%
  summarise(n = n(),
            y = sum(status == "Deceased"))

desc_residents$DAT_BFR_outbreak_homes$RR <- relative_risk(
  desc_residents$DAT_BFR_outbreak_homes$y,
  desc_residents$DAT_BFR_outbreak_homes$n - desc_residents$DAT_BFR_outbreak_homes$y,
  desc_residents$DAT_BFR_other_homes$y, 
  desc_residents$DAT_BFR_other_homes$n - desc_residents$DAT_BFR_other_homes$y)

desc_residents$DAT_BFR_outbreak_homes$RR_SE <- relative_risk_SE(
  desc_residents$DAT_BFR_outbreak_homes$y, 
  desc_residents$DAT_BFR_outbreak_homes$n - desc_residents$DAT_BFR_outbreak_homes$y,
  desc_residents$DAT_BFR_other_homes$y, 
  desc_residents$DAT_BFR_other_homes$n - desc_residents$DAT_BFR_other_homes$y)

desc_residents$DAT_BFR_outbreak_homes$RR_lower <- exp(
  log(desc_residents$DAT_BFR_outbreak_homes$RR) - 
    1.96*desc_residents$DAT_BFR_outbreak_homes$RR_SE
)
desc_residents$DAT_BFR_outbreak_homes$RR_upper <- exp(
  log(desc_residents$DAT_BFR_outbreak_homes$RR) + 
    1.96*desc_residents$DAT_BFR_outbreak_homes$RR_SE
)

@@

```{r captions}
## TABLES
table_nums <- captioner::captioner(prefix = "Table")

tab_resident_descriptives <- table_nums(
  name = "tab_resident_descriptives",
  caption = paste0("Breakdown of FSHCG private bed residents by care home type, sex, age, region and status on study exit (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
incidence_CT_residents <- table_nums(
  name = "incidence_CT_residents",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections in residents according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
supp_epi_CT_oubreak_only <- table_nums(
  name = "incidence_CT_staff",
  caption =  paste0("Incidence proportions and rates of SARS-CoV-2 infections among staff according to FSHCG manager counts (", format(date_min, "%d/%m/%y"), "-", format(date_max, "%d/%m/%y"), ")"))
tab_CoxPH_confirmed <- table_nums(
  name = "tab_CoxPH_confirmed",
  caption =  paste0("Factors associated with the hazard of become a confirmed case: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))
tab_CoxPH_death <- table_nums(
  name = "tab_CoxPH_death",
  caption =  paste0("Factors associated with all-cause mortality: hazard ratios (HR) from a Cox proportional hazards model (n=", 
  formatbm(dplyr::n_distinct(residents_overview$resident_id)),
  ")"))

@@

tab_national_descriptives <- table_nums(
  name = "tab_national_descriptives",
  caption = "Total homes, beds, contract beds, and residents within FSHCG and nationwide")

@@

**Background** Covid-19 has led to a 45% increase in mortality in care home residents in England and Wales, but the proportion of deaths directly attributable to Covid-19 is uncertain due to limited testing for infection, and lack of syndromic surveillance.

**Methods:**  Retrospective cohort study using individual-level electronic records from residents and daily care home counts of infection for `r formatbm(desc_residents$CT_total_residents)` residents and `r formatbm(desc_staff$total_staff)` staff across `r dplyr::n_distinct(residents$home_code)` care homes from the Four Seasons Health Care Group (FSHCG, `r format(date_min, "%e %B %Y")`-`r format(date_max, "%e %B %Y")`). We estimated the incidence of symptoms, confirmed infections, and deaths in residents and staff and modelled risk factors for infection and mortality in residents using Cox proportional hazards.

**Findings:** `r formatbm(desc_residents$CT_total_symptomatic)` residents developed Covid-19 symptoms (`r tbrounding(desc_residents$CT_p_symptomatic)`% [95% confidence interval: `r tbrounding(desc_residents$CT_p_symptomatic_lower)`%; `r tbrounding(desc_residents$CT_p_symptomatic_upper)`%]), while `r formatbm(desc_residents$CT_total_confirmed)` residents (`r tbrounding(desc_residents$CT_p_confirmed)`% [`r tbrounding(desc_residents$CT_p_confirmed_lower)`%; `r tbrounding(desc_residents$CT_p_confirmed_upper)`%]) and `r formatbm(desc_staff$CT_confirmed_total)` staff (`r tbrounding(desc_staff$CT_confirmed_p)`% [`r tbrounding(desc_staff$CT_confirmed_p_lower)`%; `r tbrounding(desc_staff$CT_confirmed_p_upper)`%]) had confirmed infections. The incidence of confirmed infection in residents and staff respectively was `r tbrounding(desc_residents$CT_IR_confirmed)` [`r tbrounding(desc_residents$CT_IR_confirmed_lower)`; `r tbrounding(desc_residents$CT_IR_confirmed_upper)`] and `r tbrounding(desc_staff$CT_confirmed_IR)` [`r tbrounding(desc_staff$CT_confirmed_IR_lower)`; `r tbrounding(desc_staff$CT_confirmed_IR_upper)`] per 100,000 person-days across all care homes. This includes `r desc_residents$CT_zero_case`/`r desc_residents$total_homes` homes (`r tbrounding( desc_residents$CT_zero_case/desc_residents$total_homes*100)`%) which had not reported a single confirmed infection or Covid-related death.

Among `r formatbm(desc_residents$DAT_total_residents)` private residents, `r formatbm(desc_residents$DAT_total_deaths)` all-cause deaths occurred either in homes or in hospital (`r tbrounding(desc_residents$DAT_p_death)`% [`r tbrounding(desc_residents$DAT_p_death_lower)`%; `r tbrounding(desc_residents$DAT_p_death_upper)`%]), including `r desc_residents$DAT_CFR$y` deaths of residents who had had a confirmed infection. This suggests an all-cause case-fatality rate of `r tbrounding(desc_residents$DAT_CFR_rate)`% [`r tbrounding(desc_residents$DAT_CFR_rate_lower)`%; `r tbrounding(desc_residents$DAT_CFR_rate_upper)`%], of which an estimated `r tbrounding(((desc_residents$DAT_CFR_RR - 1)/desc_residents$DAT_CFR_RR)*100)` [`r tbrounding(((desc_residents$DAT_CFR_RR_lower - 1)/desc_residents$DAT_CFR_RR_lower)*100)`; `r tbrounding(((desc_residents$DAT_CFR_RR_upper - 1)/desc_residents$DAT_CFR_RR_upper)*100)`] is attributable to confirmed SARS-CoV-2 infections. Independently, care home reports of Covid-related deaths suggest that `r desc_residents$CT_total_cdeath`/`r formatbm(desc_residents$CT_total_residents)` residents (`r tbrounding(desc_residents$CT_p_cdeath)`% [`r tbrounding(desc_residents$CT_p_cdeath_lower)`; `r tbrounding(desc_residents$CT_p_cdeath_upper)`]) died of a cause directly attributable to Covid-19. In addition to male gender and older age, confirmed asymptomatic (HR = 3·3 [1·9; 5·6]) and confirmed symptomatic (HR = 12 [10; 15]) infection were independently associated with all-cause mortality, by comparison with asymptomatic residents living in homes with no record of confirmed infections. All-cause mortality for uninfected residents was 93% higher in care homes with outbreaks compared to those without outbreaks (adjusted HR = 1·9 [1·7; 2·2]).

**Interpretation:** Covid-19 infection directly increases residents’ risk of death, but in outbreak care homes mortality is also elevated in individuals with no record of infection compared to homes without outbreaks. Preparations for future waves of the pandemic should prioritise regular testing in care homes to address case under-ascertainment, and consider the indirect effect of Covid-19 on mortality.

@@

Care home residents are particularly vulnerable to Covid-19 infection due to their advanced age and high prevalence of comorbidity,[@Zhou2020] and this is compounded by frequent exposure to infection through close contact with infected staff members, other residents and contaminated surfaces in the care home setting. In the UK, an estimated 420,000 residents lived in approximately 11,000 homes for the elderly.[@cqcdata2020] At the peak of the pandemic the number of deaths in care home residents increased threefold over the equivalent period in 2019.[@onsDR2020]

However, this is likely to underestimate of the true number of deaths in care home residents that were caused by the pandemic, since it only includes deaths in which Covid-19 was directly implicated. Death certifications in England and Wales between 28th December 2019 and 12th June 2020 suggest that two thirds of the 29,393 excess deaths recorded in care home residents (calculated relative to equivalent period in the previous year) involved Covid-19.[@onsDR2020] As part of efforts to mitigate the anticipated second wave of infection, it is critical to understand the ratio of direct to indirect deaths in care home residents. This will inform decisions around instigating lockdown measures to protect residents and staff from infection, whilst recognising the potential unintended consequences of interrupting usual care.

Disentangling direct and indirect deaths from Covid-19 requires accurate information on the incidence and prevalence of infection among staff and residents linked to individual-level clinical outcomes. There is no syndromic surveillance for infection in care homes in England, and universal testing for SARS-CoV-2 using reverse transcriptase polymerase chain reaction (RT-PCR) was not established in care homes in England until 11th May 2020, when the government launched the whole care home testing programme.[@dhscTesting2020] Prior to this, testing was only available for care home residents or staff who were admitted to hospital, or as part of Public Health England's (PHE) outbreak investigations and within a maximum of five tests per care home. Consequently national estimates of incidence and prevalence based on test results obtained prior to 11th May will substantially underestimate the burden of infection in care home residents and staff.

Outbreak investigations are useful to estimate the proportion of residents and staff who become infected during an outbreak,[@McMichael2020; @Dora2020] but they provide limited insight into the relationship between infection and mortality across the entire pandemic period, and cannot be generalised to care homes that have not reported any outbreaks. A living systematic review[@Salcher-Konrad2020] of outbreak investigations found incidence proportions ranging between 0% and 72% among residents and between 2% and 64% among staff of long-term care facilities. Mortality rates in residents with confirmed Covid-19 infection ranged from 0 to 10%, with case fatality rates between 0 and 34%, however this review did not include estimates indirect mortality.

We used administrative data from the Four Seasons Health Care Group (FSHCG), one of the UK’s largest independent provider of long-term and respite residential and nursing care, to establish syndromic surveillance for Covid-19 across `r formatbm(dplyr::n_distinct(residents$home_code))` FSHCG homes. We used this system to estimate the incidence and prevalence of suspected and confirmed Covid-19 across a national network of care homes, and to investigate the fraction of all-cause deaths attributable to confirmed infections.

@@

In 2020, the FSHCG included 9,568 residential or nursing care beds for elderly residents, representing 9% of all beds in England, Scotland and Northern Ireland (supplementary material 3). Approximately 9% of these beds were contracted to local authorities, with the remainder for privately funded residents.

FSHCG collects administrative data on their private residents through a variety of electronic systems that are primarily used for billing purposes. This includes individual-level data on the dates of care home entry and exit for residents, demographic information, and reports of infection through an incidents management system known as “Datix”. Such records do not include residents occupying beds contracted by local authorities, but home-level aggregate information was nevertheless available on all residents (`r fig_nums("fig_study_diagram", display = "cite")`).

@@

On 24 March 2020, FSHCG introduced a bespoke reporting system to monitor Covid-19 cases in each care home for all care home residents and staff (including residents in contract beds). This new system requires care homes to report daily tallies of new cases identified among staff and residents to their FSHCG regional manager, and was used to report Covid-19 deaths to the Care Quality Commission (CQC).[@CQC2020stats] Daily counts were obtained for every care home:

- new symptomatic cases (residents)
- newly confirmed infection in home (residents)
- newly confirmed infection in hospital (residents)
- deaths related to Covid-19 (residents previously confirmed or with a death attributed to SARS-CoV-2 by coroner)
- symptomatic cases (staff)
- newly confirmed cases (staff)

In addition, the number of beds currently occupied was reported every Sunday. This weekly information was used to approximate daily occupancy using linear interpolation. Aggregate data were only extracted for residents in care homes for the elderly.

Care home characteristics were inferred from organisational data. Baseline care home occupancy was computed by averaging weekly occupancy before the first Covid-19 cases (between 1 January and 1 March 2020). This was used to calculate a ratio of the baseline occupancy to the number of bedroom ratio. We also estimated a ratio of the headcount of staff employed to the number of beds.

@@

- **a residents dataset** containing one record per resident per care home, alongside residents' gender, date of birth, most recent dates of admission and discharge, date of first admission, type of stay (residential/nursing) and care (general, dementia, elderly). This dataset does not include any records for `r formatbm(sum(desc_homes[["Total contract beds"]], na.rm=T))` occupants of local authority 'block contract' beds.
- **an incidents dataset** containing `r formatbm(desc_before_exclusions$nrow_incidents)` reports filed by care home staff. Mandated report field were: resident forename, surname, care home identifier, incident date/time, and date/time of reporting. In addition, reports could record: date of birth, gender, information on Covid-19 symptoms, tests and test results, resident current location (home/hospital), and death. Incidents were classified to indicate whether the resident was symptomatic, and whether an infection was confirmed on the basis of multiple variables (supplementary material 1).

HW created a resident index file bridging the two datasets using: the first three letters of the resident's forename; the resident's full surname; and the care home identifier. Clerical review of identifiers generated in this way established that no single identifier was allocated to distinct individuals in the residents' dataset. Just `r desc_before_exclusions$missing_id` (`r tbrounding(desc_before_exclusions$missing_id_pct)`%) incident records did not contain enough information to generate an identifier and were excluded. Identifiers of a further `r desc_before_exclusions$link_fail` records (`r desc_before_exclusions$link_fail_pct`%) could not be successfully linked to the residents dataset and were excluded, after clerical review established they could not be matched with a resident. A proportion of these records may have been filed in error, while others may relate to occupants of local authority beds who are not accounted for in the residents' dataset.
